---
title: "Untitled"
author: "Yingjie"
date: "`r Sys.Date()`"
output: html_document
---


## Funnel plot for publication bias

Given our assumptions, and in the case when there is no publication bias,
 all studies would lie symmetrically around our pooled effect size (the vertical line in the middle),
 within the form of the funnel.

When *publication bias* is present, we would assume that the funnel would look asymmetrical,
 because only the small studies with a large effect size very published,
 while small studies without a significant, large effect would be missing.

We can see in the plot that while some studies have statistically significant effect
 sizes (the gray areas), others do not (white background).


### set up

```{r}
# To clear your environment 
remove(list = ls())


# # Install and load required packages (if not already installed)
# install.packages(c("meta", "metafor", "ggplot2"))
library(dplyr)
library(stringr)
library(meta)
library(metafor) # regtest()
library(ggplot2)


# Define fill colors for contour
col.contour = c("gray75", "gray85", "gray95")
```


### load data

```{r}
## load data
f <- "./data/data_ma_smd_combined.rds"
rds_list <- readRDS(f)


## loop each mental health tool
tool_selected_rct <- c('PANAS', 'POMS', 'STAI', 'SVS', 'DASS-21', 'PSS', 'ROS')
```


### plot

```{r}

for (f in names(rds_list)) {
  
  print(f)
  
  tool_ind <- basename(f) %>% gsub('ma_smd_|_.rds', '', .) 
  
  # Split into two parts
  split_string <- str_split_fixed(tool_ind, "_", 2)
  mh_tool <- split_string[1]
  ind <- split_string[2]
  
  # Check if input_string is in valid_values, otherwise stop with an error
  if (!(mh_tool %in% tool_selected_rct)) {
    stop(paste("Error: '", mh_tool, "' is not in the valid tool list! Please double-check!"))
  }
  
  ma_smd <- rds_list[[f]]
  
  
  # ========================
  # 1. Funnel Plot
  # ========================
  
  # skip now and will be included in step 5
  
  
  # ========================
  # 2. Egger’s Test (Statistical Test for Funnel Plot Asymmetry)
  # ========================
  
  metafor_data <- data.frame(
    yi = ma_smd$TE,    # Extract effect sizes (Hedges' g)
    sei = ma_smd$seTE  # Extract standard errors
  )
  egger_test <- regtest(metafor_data$yi, metafor_data$sei, model = "lm")
  # egger_test
  egger_pval <- round(egger_test$pval, 3)  # Extract p-value for annotation
  egger_plab <- ifelse(egger_pval < 0.05, 
                       'p < 0.05 (Sig. publication bias)', 
                       'p > 0.05 (No sig. publication bias)')
  egger_pcol <- ifelse(egger_pval < 0.05, 'red', 'blue')
  
  # Interpretation:
  # - If p < 0.05 → Significant bias detected
  # - If p > 0.05 → No significant bias
  
  
  
  # ========================
  # 3. Trim-and-Fill Method (Correction for Missing Studies)
  # ========================
  trimfill_result <- trimfill(ma_smd)
  # summary(trimfill_result)
  

  # ========================
  # 4. Extract data for viz
  # ========================
  ## Extract the original and adjusted effect sizes
  original_effect_size <- round(ma_smd$TE.random, 2)
  adj <- quantile(ma_smd$TE.random, 0.8)
  adjusted_effect_size <- round(trimfill_result$TE.random, 2)
  
  # Extract Key Values for Annotations
  y_upper <- max(ma_smd$seTE, na.rm = T)    # Highest standard error
  y_lower <- min(ma_smd$seTE, na.rm = T)    # Lowest standard error
  
  
  
  # ========================
  # 5. Combine both plots into one image
  # ========================

  par(mfrow = c(1, 2))  # Arrange plots side by side
  
  funnel(ma_smd, 
         main = "Original Funnel Plot", 
         # contour = c(0.9, 0.95, 0.99),
         # col.contour = col.contour, legend=TRUE,
         col = "blue", cex = 1.5)
  # # Add a legend
  # legend(x = quantile(ma_smd$TE, 0.97, na.rm = T), y = 0.01, legend = c("p < 0.1", "p < 0.05", "p < 0.01"), fill = col.contour)
  mtext(paste(mh_tool, ind, sep = ' - '), side = 3, line = 0.5, adj = 0.01, cex = 1.5)
  mtext(paste("Original Hedges' g =", original_effect_size), side = 3, line = -2, adj = 0.01, cex = 1.2)
  mtext(paste("Egger’s test:", egger_plab), side = 3, line = -4, adj = 0.02, cex = 1.2, col = egger_pcol)

  # Add Annotations to the Funnel Lines
  text(original_effect_size, y_upper, labels = "Mean Effect Size", pos = 3, cex = 1.1, col = "black")
  text(original_effect_size - adj, y_upper - 0.05, labels = "95% CI", pos = 3, cex = 1.1, col = "black")
  text(original_effect_size + adj, y_upper - 0.05, labels = "95% CI", pos = 3, cex = 1.1, col = "black")
  
  
  # Define colors with transparency
  original_color <- adjustcolor("blue", alpha.f = 0.6)  # 60% opacity
  adjusted_color <- adjustcolor("red", alpha.f = 0.4)    # 60% opacity
  
  funnel(trimfill_result, 
         main = "Trim-and-Fill Adjusted Funnel Plot",
         # col = c("red"),
         # bg = c("blue", 'red'), 
         col = ifelse(trimfill_result$studlab %in% ma_smd$studlab, original_color, adjusted_color),
         pch = ifelse(trimfill_result$studlab %in% ma_smd$studlab, 21, 16),
         cex = 1.5)
  legend("topright", legend = c("Original Studies", "Trim-and-Fill Adjusted"), 
         col = c("blue", "red"), 
         pt.bg = c("gray", 'red'), pch = 21, pt.cex = 1.5, bty = "n", cex = 1.2)
  mtext(paste(mh_tool, ind, sep = ' - '), side = 3, line = 0.5, adj = 0.01, cex = 1.5)
  mtext(paste("Adjusted Hedges' g =", adjusted_effect_size), side = 3, line = -2, adj = 0.01, cex = 1.2)
  
}

```

